---
title: "Getting average GHSL built-up and land areas for each urban area of interest"
format: html
editor_options: 
  chunk_output_type: console
---

```{r load-packages}
library(tidyverse) # CRAN v2.0.0
library(sf)        # CRAN v1.0-19
library(units)     # CRAN v0.8-5
library(stars)     # CRAN v0.6-7
library(here)      # CRAN v1.0.1
```

We could run this using the 10m resolution tiles (see `R/02_get_GHSL_buffers_highres.qmd`), but since the urban areas polygons are drawn at 1km resolution and there are GHSL `tif` files at 1 km resolution, we use these coarser versions to make calculations faster (we checked on a subset of urban areas, and results would be the same)

The `data/GHSL_1000m` folder and sub-folders is where to put `tif` files downloaded from the GHSL website (see `data/GHSL_1000m/README.md` for more details). This code only runs if these files are there. If you are checking this directly from the archived repository (which does not contain a copy of these files as they are heavy), you can skip this code and go directly to the script files that use its `csv` output, a copy of which **is** archived in this repo.

From here things are pretty straightforward, from loading the rasters as `stars` objects...

```{r stars-rasters}
stars_BUILT2020 <- read_stars(
    here("data","GHSL_1000m","BUILT-S","GHS_BUILT_S_E2020_GLOBE_R2023A_54009_1000_V1_0.tif"),
    proxy = TRUE
  )

stars_LAND <- read_stars(
    here("data","GHSL_1000m", "LAND","GHS_LAND_E2018_GLOBE_R2022A_54009_1000_V1_0.tif"),
    proxy = TRUE
  )
```

... to loading the urban area polygons...

```{r load-urban-polygons}
polygons <- read_sf( here("data","GHSL_1000m","GHSL_SMOD_UC_DUC_2020_polygons_V2_0_R2023A.gpkg"))
```

...summing built-up and land over the polygons...

```{r compute-sums-over-polygons}
sum_BUILT2020 <- aggregate(x=stars_BUILT2020,by=polygons,FUN=sum,na.rm=TRUE) |> 
  st_as_sf()|> 
  as_tibble() |> 
  select(-geometry) |> 
  rename_at(1,~"sum_BUILT2020")

sum_LAND <- aggregate(x=stars_LAND,by=polygons,FUN=sum,na.rm=TRUE) |> 
  st_as_sf()|> 
  as_tibble() |> 
  select(-geometry) |> 
  rename_at(1,~"sum_LAND")
```

... and finally using these to calculate % of built-up and land:

```{r summarise-results}
urban_summary <- cbind(sum_BUILT2020,sum_LAND) |> 
  mutate(urban_area=polygons$urban_area, .before = 1) |> 
  mutate(AREA=as.numeric(st_area(polygons))) |> 
  mutate(area_km2_classic = AREA/ (1000*1000)) |> 
  mutate(
    mean_pBUILT_classic = sum_BUILT2020/AREA,
    mean_pBUILT = sum_BUILT2020/sum_LAND,
    mean_pLAND = sum_LAND/AREA
  ) |> 
  mutate(area_LAND_km2 = mean_pLAND * area_km2_classic) |> 
  select(urban_area,
         area_km2_classic,
         area_LAND_km2,
         mean_pLAND,
         mean_pBUILT,
         mean_pBUILT_classic
         )

urban_summary
```


```{r write-csv}
write_csv(urban_summary, here("data","GHSL_urban_summary.csv"))
```
